A sheaf--function dictionary for algebraic tori over local fields

### 2013-09-19 shanghai.tex
I got a copy of the paper by Deligne and made a reference to it slightly more specific.
Adding QC-schemes, breaking long lines, fixing typos

\begin{abstract}
  We introduce and study the category of quasicharacter sheaves on smooth,
  commutative group schemes $G$ locally of finite type over finite fields
  $\Fq$. Assuming that the geometric component group of $G$
  is finitely-generated, we show that the group of isomorphism classes
  of quasicharacter sheaves on $G$ is canonically isomorphic to the group
  of characters of $G(\Fq)$. We use this result to exhibit an
  isomorphism between quasicharacters of $T(K)$ and isomorphism
  classes of quasicharacter sheaves on the Greenberg transform of the Néron
  model of $T$, where $K$ is any non-archimdean local field and $T$ is
  any algebraic torus over $K$.
\end{abstract}

@@

Throughout this paper, $G$ is a smooth, commutative group scheme locally of finite type
over a finite field $\Fq$, and $m : \GxG\to G$ is its multiplication morphism.

@@

The category $\QC(G)$ of quasicharacter sheaves on $G$ is a rigid symmetric monoidal category
\cite{etingof:09a}*{\S~1.10} under the tensor product
$\qcs{L} \otimes \qcs{L'} \ceq (\gqcs{L}\otimes\gqcs{L'}, \mu\otimes\mu', \phi\otimes \phi')$,
with duals given by applying the sheaf hom functor 
$\sheafHom(\ - \ ,\EE)$.
The category of quasicharacter sheaves on $G$ is not abelian, so $\QC(G)$
is not a tensor category in the sense of \cite{deligne:02a}*{0.1}.
This rigid monoidal category structure for $\QC(G)$ puts a group
structure on the set $\QCiso{G}$ of isomorphism classes in $\QC(G)$.
We now begin developing the tools to relate $\QCiso{G}$ to $\Hom(G(\Fq), \EEx)$.

@@

The isomorphism $\mu : \bm^* \gqcs{L} \to \LxL$ together with property~\ref{CS.2} guarantee
that the function $\trFrob{\qcs{L}} : G(\Fq)\to \EEx$ is a group homomorphism.
It is also clear that $\trFrob{\qcs{L}\otimes \qcs{L'}} = \trFrob{\qcs{L}} \cdot \trFrob{\qcs{L'}}$
(pointwise multiplication of functions) and $\trFrob{\qcs{L}} = \trFrob{\qcs{L'}}$
if $\qcs{L} \iso \qcs{L'}$ in $\QC(G)$. In this way we obtain a group homomorphism
\[
\cdef{\trFrob{G}} : \QCiso{G} \longrightarrow G(\Fq)^*,
\qquad \text{defined by}\qquad \qcs{L} \mapsto \trFrob{\qcs{L}}.
\]
This isomorphism is functorial in the following sense:

\begin{proposition}\label{prop:functorialG}
Trace of Frobenius $\QCiso{G} \to G(\Fq)$ is a natural transformation 
between the two additive functors 
\begin{enumerate}
\item[$F_1$:] $G \mapsto \QCiso{G}$
\item[$F_2$:] $G \mapsto G(\Fq)^*$
\end{enumerate}
from the category of smooth commutative
group schemes locally of finite type over $\Fq$ to the category of abelian groups.
\end{proposition}

@@

%\David{insert some commentary on this proposition: it will be very useful to us}
As a step toward the proof of Proposition~\ref{prop:functorialG} we show that every morphism  $H \to G$ of smooth group schemes locally of finite type over $\Fq$ induces a monoidal functor from quasicharacter sheaves on $G$ to quasicharacters on $H$. This simple but important result will also play a role in the proof of many other propositions in this paper.

\begin{proposition}\label{prop:pullback}
  If $f : H\to G$ is a morphism of commutative group schemes locally of finite type over $\Fq$ then
  $(\gqcs{L},\mu,\phi) \mapsto (\bf^*\gqcs{L},(\bfxf)^*\mu,\bf^*F)$
  defines a monoidal functor $f^* : \QC(G) \to \QC(H)$ such that
  \[
  \begin{tikzcd}[row sep=20, column sep=30]
   \QCiso{G} \rar{f^*} \arrow[swap]{d}{\trFrob{G}} & \QCiso{H} \dar{\trFrob{H}} \\
   G(\Fq)^* \rar & H(\Fq)^*
  \end{tikzcd}
  \]
  is a commutative diagram of groups, where the lower homomorphism is
  dual to $f : H(\Fq)\to G(\Fq)$.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{proposition}

@@

Consider the short exact sequence in the category of smooth group
schemes, locally of finite type over $\Fq$, defining the component
group scheme for $G$:
\begin{equation}\label{eq:pi0}
\begin{tikzcd}
0 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
\end{tikzcd}
\end{equation}
Since $\pi_0(G)$ is an commutative group scheme, smooth and locally of finite type over $\Fq$, 
Proposition~\ref{prop:pullback} implies that \eqref{eq:pi0} defines a sequence of functors
\[
\begin{tikzcd}
\QC(0) \rar & \QC(\pi_0(G)) \rar & \QC(G) \rar & \QC(G^0) \rar & \QC(0)
\end{tikzcd}
\]
which induces, upon passage to isomorphism classes of objects, a sequence of group homomorphisms 
\[
\begin{tikzcd}
\QCiso{0} \rar & \QCiso{\pi_0(G)} \rar & \QCiso{G} \rar & \QCiso{G^0} \rar & \QCiso{0}.
\end{tikzcd}
\]
It is easy to see that $\QCiso{0}$ is trivial. In Sections~\ref{sec:connected} and  we will find $\QCiso{G^0}$
and $\QCiso{\pi_0(G)}$ under the hypothesis that $\pi_0(G)$ is geometrically finitely generated; 
the snake lemma will then allow us to show
\[
\QCiso{G} \iso G(\Fq)^*,%\Hom_\text{grp}(G(\Fq),\EE^\times),
\]
canonically.
%\section{Products} 
We end this section with a simple result on products of quasicharacter sheaves,
which is needed to show that the isomorphism above is functorial.

@@

Finally, let $G'$ be a smooth commutative group scheme locally of finite type over $k'$;
we explain how to geometrize the canonical isomorphism between characters of $G'(k')$ and of $(\Res_{k'/k}G')(k)$.
The base change $(\Res_{k'/k}G')_{k'}$ of $\Res_{k'/k}G'$ to $k'$
decomposes into a product of copies of $G'$, indexed by elements of $\Gal(k'/k)$:
\[
(\Res_{k'/k}G')_{k'} \cong \prod_{\Gal(k'/k)} G'.
\]
Since products and coproducts agree for group schemes we have a natural inclusion of $k'$-schemes 
\[
G' \hookrightarrow (\Res_{k'/k}G')_{k'}
\]
mapping $G'$ into the summand corresponding to $1 \in \Gal(k'/k)$.  Composing pullback
along this map with $\nu_{k'/k}$ from Proposition~\ref{prop:basechange} yields a functor
\[
\QC(\Res_{k'/k}G') \to \QC(G').
\]

@@

In principle, something similar is possible in the more general 
context of this paper, where $G$ is a smooth group scheme 
locally of finite type over $\Fq$, to which we now return. 
In this generality, neither of the full subcategories $\QCf(G) \subset \QCb(G) \subset \QC(G)$ are equivalences;
see the end of Section~\ref{sec:snake}.
Nevertheless, every quasicharacter sheaf $(\gqcs{L},\mu,\phi)$ on $G$ 
determines a Weil sheaf $(\gqcs{L},\phi)$ on $G$. 
So, if we choose a geometric point ${\bar g}_x$ in each component $\bG^x$ of $G$, then this choice can be used to convert 
the Weil sheaf $(\gqcs{L},\phi)$ into an $\ell$-adic character of $\prod_{x\in \pi_0(\bG)}\W(\bG^x, {\bar g}_x)$. 
However, it proved cumbersome to manage various 
choices of families of base points and track the action 
of $\Gal(\bFq/\Fq)$ on this (generally) infinite product of 
Weil groups and difficult to identify the relevant quotient of this group.
Viewed in this light, Definition~\ref{def:QC} is comparatively simple.

@@

\'Etale group schemes form a counterpoint to connected groups,
since the component group of a group scheme locally of finite type over $\Fq$
is an \'etale group scheme (locally of finite type) over $\Fq$
\cite{vdGeer-Moonen:AbelianVarieties}*{III, \S 4}.
The functor $G \mapsto G(\bFq)$ defines an equivalence of categories
between the category of \'etale group schemes over $\Fq$ and the category of groups equipped
with an action of $\Gal(\bFq/\Fq)$, continuous for the discrete topology on the group.
Under this equivalence, a quasicharacter sheaf on $G$ is given by the data of:
\begin{enumerate}
 \labitem{(cs.0)}{cs.0} an indexed set of one-dimensional
  $\EE$-vector spaces $\gqcs{L}_x$, as $x$ runs over
  $G(\bFq)$;

@@

The main result of Part~\ref{part1} of this paper, 
the sheaf--function dictionary for $G$ Theorem~\ref{thm:snake}, has a hypothesis: 
the smooth commutative group scheme locally of finite type must have the 
property that the geometric component group of $G$ is finitely generated as a group. 
The genesis of that hypothesis is Lemma~\ref{lemma:section},
as we use this hypothesis to show the existence of global sections of $\gqcs{L}$ satisfying \eqref{section} for $\mu$.
Remarkably, Lemma~\ref{lemma:section} is the only crux for this hypothesis;
that is, if $G$ is any smooth connected group scheme locally of finite type over $\Fq$
and if each quasicharacter sheaf on the \'etale group scheme $\pi_0(G)$ admits a global section 
satisfying \eqref{section}, then the conclusion of Theorem~\ref{thm:snake} holds for $G$,
so there is a sheaf--function dictionary for $G$.

@@

\begin{proof}
  Since $G$ is a smooth group scheme locally of finite type over $\Fq$, its
  identity component $G^0$ of $G$ is a connected smooth, 
  geometrically reduced group scheme of finite type over $\Fq$, by
   \cite{vdGeer-Moonen:AbelianVarieties}*{3.17}. 
	Since $\Fq$ is a finite field, $G^0$ is actually reduced over $\Fq$;
	see \cite{stacks-project}*{\href{http://stacks.math.columbia.edu/tag/035X}{Tag 035X}}.
	It follows that $G^0$ is a connected algebraic group.
    By Proposition~\ref{prop:connected}, every
  quasicharacter sheaf on $G^0$ is isomorphic to a 
  finite quasicharacter sheaf on $G^0$, 
  so to prove the proposition it suffices to show that every 
  finite quasicharacter sheaf on $G^0$ extends to a quasicharacter sheaf on $G$.

@@

We saw, in Proposition~\ref{prop:functorialG}, that the trace of Frobenius
$\trFrob{G} : \QCiso{G} \to G(\Fq)^*$ is a functorial group homomorphism.
Our main result regarding quasicharacter sheaves, below, shows that
trace of Frobenius is an isomorphism in the case that we care about.

%\input{snake}
\begin{theorem}\label{thm:snake}
  If $G$ is a smooth commutative group scheme locally of finite type over
  $\Fq$ and if the abelian group $\pi_0(G)(\bFq)$ (the geometric
  component group) is finitely generated then the trace of Frobenius
  \[
  \trFrob{G} : \QCiso{G} \to G(\Fq)^*
  \]
  is an isomorphism of groups.
\end{theorem}

\begin{proof}
  Recall the short exact sequence \eqref{eq:pi0} in the category of smooth group
  schemes, locally of finite type over $\Fq$, defining the component
  group scheme for $G$:
  \[
  \begin{tikzcd}
    1 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
  \end{tikzcd}
  \]
  Let
  \[
  \begin{tikzcd}[row sep=30]
    {}& \ker \trFrob{\pi_0(G)} \dar & \arrow[dashed]{d} \ker \trFrob{G} & \ker \trFrob{G^0} \dar & \\
    0 \rar & \QCiso{\pi_0(G)} \rar \dar{\trFrob{\pi_0(G)}}
    & \QCiso{G} \rar \dar{\trFrob{G}} & \QCiso{G^0} \rar \dar{\trFrob{G^0}} & 0\\
    0 \rar & \pi_0(G)(\Fq)^* \rar \dar
    & \arrow[dashed]{d} G(\Fq)^* \rar & G^0(\Fq)^* \rar \dar & 0\\
    & \coker \trFrob{\pi_0(G)} & \coker \trFrob{G} &  \coker \trFrob{G^0} &
  \end{tikzcd}
  \]
  be the commutative diagram of abelian groups obtained by applying
  Proposition~\ref{prop:pullback} to \eqref{eq:pi0}.  Since $k$ is perfect
  \cite{stacks-project}*{\href{http://stacks.math.columbia.edu/tag/035X}{Tag 035X}} implies that
  \David{Need another reference that an algebraic group over an algebraically closed field is reduced}
  $G$ is reduced.  Since $G^0$ is a finite type
  \cite{vdGeer-Moonen:AbelianVarieties}*{3.17}, it is thus a connected algebraic group and 
  Proposition~\ref{prop:connected} implies that $\ker \trFrob{G^0} =0$ and $\coker \trFrob{G^0}=0$.
  By Propositions~\ref{prop:etale} and \ref{prop:X}, $\ker \trFrob{\pi_0(G)}=0$
  and $\coker \trFrob{\pi_0(G)}=0$.

@@

We will also need the Greenberg functor, 
introduced in \cite{greenberg:61} and \cite{greenberg:63a}, 
as a functor from the category of schemes locally of finite type 
over an Artinian local ring with perfect residue field $k$ 
to the category of schemes locally of finite type over $k$. 
This functor respects open immersions, closed immersions, 
etale morphisms, smooth morphisms and geometric components.
Besides Greenberg's original papers, see \cite{demazure-gabriel:GroupesAlgebriques}*{V, \S 4, no. 1}, \cite{bosch-lutkebohmert-reynaud:NeronModels}*{Ch. 9, \S 6} 
and the first few sections of \cite{stasinski:12a} 
for background on the Greenberg functor.

@@

Some fundamental properties of $f_n : \GN{T}_{n+1} \to \GN{T}_n$ are:
 \begin{enumerate}
   \labitem{(GN.0)}{points} $\GN{T}_n(\Fq) = \TT(\OK/\pK^n)$, canonically.
  \labitem{(GN.1)}{smooth} 
  $\GN{T}_n$ is a smooth commutative group scheme locally of finite type over $\Fq$ 
  and the component group of $\GN{T}_n$ is geometrically finitely generated. 
  Note that $\underline{T}$ is a smooth group scheme over $\OK$, by construction. 
  Since smoothness is preserved under base change, 
  $\TT_n \ceq \underline{T} \times_\Spec{\OK} \Spec{\OK/\pK^n}$ 
  is a smooth group scheme over an Artinian ring with perfect residue field. 
  The Greenberg functor preserves smoothness, 
  so $\GN{T}_n$ is a smooth group scheme locally of finite type over $\Fq$. 
  The Greenberg functor takes the geometric identity component of $\TT_n$ 
  to the geometric identity component of $\GN{T}_n$;
  in fact, $\pi_0(\GN{T}_n) = \pi_0(\TT) \times_{\Spec{\OK}} \Spec{\Fq}$ and
  this group is finitely generated by \cite{xarles:93a}*{2.18}.
  \labitem{(GN.2)}{truncation}
  The morphism $f_n : \GN{T}_{n+1} \to \GN{T}_n$ is a surjective morphism of group schemes over $\Fq$; 
  see  \cite{bosch-lutkebohmert-reynaud:NeronModels}*{p. 276}
  and  \cite{stasinski:12a}*{\S 4}.
  We remark that $f_n : \GN{T}_{n+1} \to \GN{T}_n$ is not induced from a morphism of Néron models.
\end{enumerate}

@@

The techniques in this paper apply to all non-archimdean local fields $K$ with residue field $k$,
and to all algebraic tori $T$ over $K$, without placing any restrictions on the characteristic of $K$
or on the ramification of $T$. Because of this, it is natural to ask if we can compare
quasicharacter sheaves on $T$ and on $T'$ under conditions when there is a
natural comparison between characters of $T(K)$ and of $T'(K')$.
This comparison of has been made precise in \cite{chai-yu:01a}, \David{more specific citation?}\Clifton{In this case, maybe not necessary.}
building on ideas of \cite{deligne:84a}.

>>>>

\begin{abstract}
  We introduce and study the category of quasicharacter sheaves on smooth,
  commutative group schemes $G$ locally of finite type over finite fields
  $\Fq$. Assuming that the geometric component group of $G$
  is finitely-generated, we show that the group of isomorphism classes
  of quasicharacter sheaves on $G$ is canonically isomorphic to the group
  of characters of $G(\Fq)$. We use this result to exhibit an
  isomorphism between quasicharacters of $T(K)$ and isomorphism
  classes of quasicharacter sheaves on the Greenberg transform of the Néron
  model of $T$, where $K$ is any non-archimedean local field and $T$ is
  any algebraic torus over $K$.
\end{abstract}

@@

Throughout this paper, $G$ is a smooth, commutative group scheme locally of finite type
over a finite field $\Fq$, and $m : \GxG\to G$ is its multiplication morphism.  For the sake of brevity, we
define a \emph{$\QC$-scheme} to be such a $G$ with the additional property that the
geometric component group of $G$ is finitely generated.  This last assumption is only
needed in the proof of Lemma \ref{lemma:section}.

@@

The category $\QC(G)$ of quasicharacter sheaves on $G$ is a rigid symmetric monoidal category
\cite{etingof:09a}*{\S~1.10} under the tensor product
$\qcs{L} \otimes \qcs{L'} \ceq (\gqcs{L}\otimes\gqcs{L'}, \mu\otimes\mu', \phi\otimes \phi')$,
with duals given by applying the sheaf hom functor 
$\sheafHom(\ - \ ,\EE)$.
The category of quasicharacter sheaves on $G$ is not abelian, so $\QC(G)$
is not a tensor category in the sense of \cite{deligne:02a}*{0.1}.
This rigid monoidal category structure for $\QC(G)$ puts a group
structure on the set $\QCiso{G}$ of isomorphism classes in $\QC(G)$.
We now begin developing the tools to relate $\QCiso{G}$ to $G(\Fq)^*$.

@@

The isomorphism $\mu : \bm^* \gqcs{L} \to \LxL$ together with property~\ref{CS.2} guarantee
that the function $\trFrob{\qcs{L}} : G(\Fq)\to \EEx$ is a group homomorphism.  If we write
$\trFrob{\qcs{L}} \cdot \trFrob{\qcs{L'}}$ for pointwise multiplication, then 
 $\trFrob{\qcs{L}\otimes \qcs{L'}} = \trFrob{\qcs{L}} \cdot \trFrob{\qcs{L'}}$.
Since $\qcs{L} \iso \qcs{L'}$ implies $\trFrob{\qcs{L}} = \trFrob{\qcs{L'}}$,
we obtain a group homomorphism
\[
\cdef{\trFrob{G}} : \QCiso{G} \longrightarrow G(\Fq)^*,
\qquad \text{defined by}\qquad \qcs{L} \mapsto \trFrob{\qcs{L}}.
\]
This homomorphism is functorial in the following sense:

\begin{proposition}\label{prop:functorialG}
Trace of Frobenius $\QCiso{G} \to G(\Fq)^*$ is a natural transformation 
between the two additive functors 
\begin{enumerate}
\item[$F_1$:] $G \mapsto \QCiso{G}$
\item[$F_2$:] $G \mapsto G(\Fq)^*$
\end{enumerate}
from the category of $\QC$-schemes to the category of abelian groups.
\end{proposition}

@@

%\David{insert some commentary on this proposition: it will be very useful to us}
As a step toward the proof of Proposition~\ref{prop:functorialG} we show that every
morphism  $H \to G$ of $\QC$-schemes induces a monoidal functor from quasicharacter
sheaves on $G$ to quasicharacters on $H$. This simple but important result will also
play a role in the proof of many other propositions in this paper.

\begin{proposition}\label{prop:pullback}
  If $f : H\to G$ is a morphism of $\QC$-schemes then
  $(\gqcs{L},\mu,\phi) \mapsto (\bf^*\gqcs{L},(\bfxf)^*\mu,\bf^*F)$
  defines a monoidal functor $f^* : \QC(G) \to \QC(H)$ such that
  \[
  \begin{tikzcd}[row sep=20, column sep=30]
   \QCiso{G} \rar{f^*} \arrow[swap]{d}{\trFrob{G}} & \QCiso{H} \dar{\trFrob{H}} \\
   G(\Fq)^* \rar & H(\Fq)^*
  \end{tikzcd}
  \]
  is a commutative diagram of groups, where the lower homomorphism is
  dual to $f : H(\Fq)\to G(\Fq)$.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{proposition}

@@

Consider the short exact sequence in the category of $\QC$-schemes
defining the component group scheme for $G$:
\begin{equation}\label{eq:pi0}
\begin{tikzcd}
0 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
\end{tikzcd}
\end{equation}
Since $\pi_0(G)$ is also a $\QC$-scheme, 
Proposition~\ref{prop:pullback} implies that \eqref{eq:pi0} defines a sequence of functors
\[
\begin{tikzcd}
\QC(0) \rar & \QC(\pi_0(G)) \rar & \QC(G) \rar & \QC(G^0) \rar & \QC(0)
\end{tikzcd}
\]
which induces, upon passage to isomorphism classes of objects, a sequence of group homomorphisms 
\[
\begin{tikzcd}
\QCiso{0} \rar & \QCiso{\pi_0(G)} \rar & \QCiso{G} \rar & \QCiso{G^0} \rar & \QCiso{0}.
\end{tikzcd}
\]
It is easy to see that $\QCiso{0}$ is trivial. In Sections~\ref{sec:connected} and  we will find $\QCiso{G^0}$
and $\QCiso{\pi_0(G)}$ under the hypothesis that $\pi_0(G)$ is geometrically finitely generated; 
the snake lemma will then allow us to show
\[
\QCiso{G} \iso G(\Fq)^*,%\Hom_\text{grp}(G(\Fq),\EE^\times),
\]
canonically.
%\section{Products} 
We end this section with a simple result on products of quasicharacter sheaves,
which is needed to show that the isomorphism above is functorial.

@@

Finally, let $G'$ be a $\QC$-scheme over $k'$;
we explain how to geometrize the canonical isomorphism between characters of $G'(k')$ and of $(\Res_{k'/k}G')(k)$.
The base change $(\Res_{k'/k}G')_{k'}$ of $\Res_{k'/k}G'$ to $k'$
decomposes into a product of copies of $G'$, indexed by elements of $\Gal(k'/k)$:
\[
(\Res_{k'/k}G')_{k'} \cong \prod_{\Gal(k'/k)} G'.
\]
Since products and coproducts agree for group schemes we have a natural inclusion of $k'$-schemes 
\[
G' \hookrightarrow (\Res_{k'/k}G')_{k'}
\]
mapping $G'$ into the summand corresponding to $1 \in \Gal(k'/k)$.  Composing pullback
along this map with $\nu_{k'/k}$ from Proposition~\ref{prop:basechange} yields a functor
\[
\QC(\Res_{k'/k}G') \to \QC(G').
\]

@@

In principle, something similar is possible in the more general 
context of this paper, where $G$ is a $\QC$-scheme, to which we now return. 
In this generality, neither of the full subcategories $\QCf(G) \subset \QCb(G) \subset \QC(G)$ are equivalences;
see the end of Section~\ref{sec:snake}.
Nevertheless, every quasicharacter sheaf $(\gqcs{L},\mu,\phi)$ on $G$ 
determines a Weil sheaf $(\gqcs{L},\phi)$ on $G$. 
So, if we choose a geometric point ${\bar g}_x$ in each component $\bG^x$ of $G$, then this choice can be used to convert 
the Weil sheaf $(\gqcs{L},\phi)$ into an $\ell$-adic character of $\prod_{x\in \pi_0(\bG)}\W(\bG^x, {\bar g}_x)$. 
However, it proved cumbersome to manage various 
choices of families of base points and track the action 
of $\Gal(\bFq/\Fq)$ on this (generally) infinite product of 
Weil groups and difficult to identify the relevant quotient of this group.
Viewed in this light, Definition~\ref{def:QC} is comparatively simple.

@@

\'Etale group schemes form a counterpoint to connected groups,
since the component group of a $\QC$-scheme
is an \'etale group scheme, locally of finite type over $\Fq$
\cite{vdGeer-Moonen:AbelianVarieties}*{III, \S 4}.
The functor $G \mapsto G(\bFq)$ defines an equivalence of categories
between the category of \'etale group schemes over $\Fq$ and the category of groups equipped
with an action of $\Gal(\bFq/\Fq)$, continuous for the discrete topology on the group.
Under this equivalence, a quasicharacter sheaf on $G$ is given by the data of:
\begin{enumerate}
 \labitem{(cs.0)}{cs.0} an indexed set of one-dimensional
  $\EE$-vector spaces $\gqcs{L}_x$, as $x$ runs over
  $G(\bFq)$;

@@

We note that this lemma is the only ingredient of the proof of Theorem~\ref{thm:snake}
that forces $\QC$-schemes to have finitely generated geometric component group.  In fact,
if $G$ is any smooth connected group scheme locally of finite type over $\Fq$
and if each quasicharacter sheaf on the \'etale group scheme $\pi_0(G)$ admits a global section 
satisfying \eqref{section}, then the conclusion of Theorem~\ref{thm:snake} holds for $G$.

@@

\begin{proof}
  Since $G$ is a smooth group scheme locally of finite type over $\Fq$, its
  identity component $G^0$ of $G$ is a connected smooth, 
  geometrically reduced group scheme of finite type over $\Fq$
   \cite{vdGeer-Moonen:AbelianVarieties}*{3.17}. 
	Since $\Fq$ is a finite field, $G^0$ is actually reduced over $\Fq$;
	see \cite{stacks-project}*{\href{http://stacks.math.columbia.edu/tag/035X}{Tag 035X}}.
	It follows that $G^0$ is a connected algebraic group.
    By Proposition~\ref{prop:connected}, every
  quasicharacter sheaf on $G^0$ is isomorphic to a 
  finite quasicharacter sheaf on $G^0$, 
  so to prove the proposition it suffices to show that every 
  finite quasicharacter sheaf on $G^0$ extends to a quasicharacter sheaf on $G$.

@@

We saw, in Proposition~\ref{prop:functorialG}, that the trace of Frobenius
$\trFrob{G} : \QCiso{G} \to G(\Fq)^*$ is a functorial group homomorphism.
Our main result regarding quasicharacter sheaves, below, shows that
trace of Frobenius is an isomorphism for $\QC$-schemes.

%\input{snake}
\begin{theorem}\label{thm:snake}
  If $G$ is a smooth commutative group scheme locally of finite type over
  $\Fq$ and if the abelian group $\pi_0(G)(\bFq)$
  is finitely generated then the trace of Frobenius
  \[
  \trFrob{G} : \QCiso{G} \to G(\Fq)^*
  \]
  is an isomorphism of groups.
\end{theorem}

\begin{proof}
  Recall the short exact sequence \eqref{eq:pi0} in the category of $\QC$-schemes
  defining the component group scheme for $G$:
  \[
  \begin{tikzcd}
    1 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
  \end{tikzcd}
  \]
  Let
  \[
  \begin{tikzcd}[row sep=30]
    {}& \ker \trFrob{\pi_0(G)} \dar & \arrow[dashed]{d} \ker \trFrob{G} & \ker \trFrob{G^0} \dar & \\
    0 \rar & \QCiso{\pi_0(G)} \rar \dar{\trFrob{\pi_0(G)}}
    & \QCiso{G} \rar \dar{\trFrob{G}} & \QCiso{G^0} \rar \dar{\trFrob{G^0}} & 0\\
    0 \rar & \pi_0(G)(\Fq)^* \rar \dar
    & \arrow[dashed]{d} G(\Fq)^* \rar & G^0(\Fq)^* \rar \dar & 0\\
    & \coker \trFrob{\pi_0(G)} & \coker \trFrob{G} &  \coker \trFrob{G^0} &
  \end{tikzcd}
  \]
  be the commutative diagram of abelian groups obtained by applying
  Proposition~\ref{prop:pullback} to \eqref{eq:pi0}.  Since $k$ is perfect
  \cite{stacks-project}*{\href{http://stacks.math.columbia.edu/tag/035X}{Tag 035X}} implies that
  \David{Need another reference that an algebraic group over an algebraically closed field is reduced}
  $G$ is reduced.  Since $G^0$ is a finite type
  \cite{vdGeer-Moonen:AbelianVarieties}*{3.17}, it is thus a connected algebraic group and 
  Proposition~\ref{prop:connected} implies that $\ker \trFrob{G^0} =0$ and $\coker \trFrob{G^0}=0$.
  By Propositions~\ref{prop:etale} and \ref{prop:X}, $\ker \trFrob{\pi_0(G)}=0$
  and $\coker \trFrob{\pi_0(G)}=0$.

@@

We will also need the Greenberg functor, 
introduced in \cite{greenberg:61} and \cite{greenberg:63a}, 
as a functor from the category of schemes locally of finite type 
over an Artinian local ring with perfect residue field $k$ 
to the category of schemes locally of finite type over $k$. 
This functor respects open immersions, closed immersions, 
etale morphisms, smooth morphisms and geometric components.
Besides Greenberg's original papers, see
\citelist{\cite{demazure-gabriel:GroupesAlgebriques}*{V, \S 4, no. 1} \cite{bosch-lutkebohmert-reynaud:NeronModels}*{Ch. 9, \S 6}}
and the first few sections of \cite{stasinski:12a} 
for background on the Greenberg functor.

@@

Some fundamental properties of $f_n : \GN{T}_{n+1} \to \GN{T}_n$ are:
 \begin{enumerate}
   \labitem{(GN.0)}{points} $\GN{T}_n(\Fq) = \TT(\OK/\pK^n)$, canonically.
  \labitem{(GN.1)}{smooth} 
  $\GN{T}_n$ is a $\QC$-scheme. 
  Note that $\underline{T}$ is a smooth group scheme over $\OK$, by construction. 
  Since smoothness is preserved under base change, 
  $\TT_n \ceq \underline{T} \times_\Spec{\OK} \Spec{\OK/\pK^n}$ 
  is a smooth group scheme over an Artinian ring with perfect residue field. 
  The Greenberg functor preserves smoothness, 
  so $\GN{T}_n$ is a smooth group scheme locally of finite type over $\Fq$. 
  The Greenberg functor takes the geometric identity component of $\TT_n$ 
  to the geometric identity component of $\GN{T}_n$;
  in fact, $\pi_0(\GN{T}_n) = \pi_0(\TT) \times_{\Spec{\OK}} \Spec{\Fq}$ and
  this group is finitely generated by \cite{xarles:93a}*{2.18}.
  \labitem{(GN.2)}{truncation}
  The morphism $f_n : \GN{T}_{n+1} \to \GN{T}_n$ is a surjective morphism of group schemes over $\Fq$; 
  see  \cite{bosch-lutkebohmert-reynaud:NeronModels}*{p. 276}
  and  \cite{stasinski:12a}*{\S 4}.
  We remark that $f_n : \GN{T}_{n+1} \to \GN{T}_n$ is not induced from a morphism of Néron models.
\end{enumerate}

@@

The techniques in this paper apply to all non-archimdean local fields $K$ with residue field $k$,
and to all algebraic tori $T$ over $K$, without placing any restrictions on the characteristic of $K$
or on the ramification of $T$. Because of this, it is natural to ask if we can compare
quasicharacter sheaves on $T$ and on $T'$ under conditions when there is a
natural comparison between characters of $T(K)$ and of $T'(K')$.
This comparison of has been made precise in \cite{chai-yu:01a}, 
building on ideas in \cite{deligne:84a}*{\S 1}.
